Perturbing SU($n$)-symmetric interactions

### 2020-02-01 collective_sun/sun_perturbation.tex
expressed triple product in terms of collective operators

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; and the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
we can write out explicitly
\begin{align}
  \diagram{two_body_2}
  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
\end{align}
and, using the fact that $v_{pp}=w_{pp}=0$,
\begin{align}
  \diagram{two_body_1}
  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
\end{align}
Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
some more work we can work out that
\begin{align}
  \diagram{two_body_0}
  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the fully symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_6},
  &&
  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
  + \diagram{triple_2_2} + \diagram{triple_2_3},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles implicitly contains a sum over
all label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
  \label{eq:trible_uvw}
\end{align}
\begin{align}
  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
  \label{eq:trible_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:trible_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:trible_uvw_3}.

>>>>

Calculating the second-order effective Hamiltonian $H_2^{(2)}$ induced
on the fully symmetric manifold $\M_0$ by the perturbation $\V_2$ in
\eqref{eq:perturbations} requires us to simplify
\begin{multline}
  \P_0 \sp{\sum_{k<\ell} \sum_{p<q}
    v_{k\ell} w_{pq} O_{k\ell} Q_{pq}} \P_0 \\
  \EQFS \diagram{two_body_0} O_{1,2} Q_{3,4}
  + \diagram{two_body_1} O_{1,2} Q_{1,3}
  + \diagram{two_body_2} O_{1,2} Q_{1,2},
  \label{eq:POQP_start}
\end{multline}
where $k,\ell,p,q$ index individual spins; $O$ and $Q$ are
permutationally-symmetric two-spin operators; and the coefficients
$v_{k\ell},w_{pq}$ satisfy \eqref{eq:cond_2} with eigenvalue
$\Delta\ne0$, which implies that they are mean-zero.  Defining
\begin{align}
  \v{\m m} \equiv \sum_{p<q} m_{pq} \ket{\set{p,q}},
  &&
  \v m \equiv \sum_p m_p \ket{p},
  &&
  m_p \equiv \sum_q m_{pq},
\end{align}
for $m\in\set{v,w}$, we can write out explicitly
\begin{align}
  \diagram{two_body_2}
  \equiv \sum_{p<q} v_{pq} w_{pq} = \v{\m v} \c\v{\m w},
\end{align}
and, using the fact that $v_{pp}=w_{pp}=0$,
\begin{align}
  \diagram{two_body_1}
  \equiv \sum_{\substack{k,p,q\\k\ne p\ne q\ne k}} v_{kp} w_{kq}
  = \sum_{k,p} v_{kp} \sum_{q\ne p} w_{kq}
  = \sum_{k,p} v_{kp} \p{w_{k} - w_{kp}}
  = \v v \c \v w - 2 \v{\m v} \c \v{\m w}.
\end{align}
Using additionally the fact that $v_{pq},w_{pq}$ are mean-zero, with
some more work we can work out that
\begin{align}
  \diagram{two_body_0}
  \equiv \sum_{k<\ell} \sum_{\substack{p<q\\p,q\notin\set{k,\ell}}}
  v_{k\ell} w_{pq} = -\v v \c \v w + \v{\m v} \c \v{\m w}.
\end{align}
Altogether, we thus have that
\begin{multline}
  \sum_{k<\ell} \sum_{p<q} v_{k\ell} w_{pq} O_{k\ell} Q_{pq} \\
  \EQFS - \p{\v v \c \v w - \v{\m v} \c \v{\m w}} O_{1,2} Q_{3,4}
  + \p{\v v\c\v w - 2 \v{\m v} \c \v{\m w}} O_{1,2} Q_{1,3}
  + \v{\m v} \c \v{\m w}\, O_{1,2} Q_{1,2}.
  \label{eq:POQP}
\end{multline}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the fully symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_6},
  &&
  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
  + \diagram{triple_2_2} + \diagram{triple_2_3},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles implicitly contains a sum over
all label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
can write \eqref{eq:triple_multi} in terms of collective operators as
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}
